\chapter*{Acknowledgements}
\lettrine{I}{} had the luck and pleasure to be surrounded by wonderful people during my Ph.D. career. Most of them do not even know that without them the reader would not be keeping this thesis in her or his hands.

First and foremost, I would like to express my immense gratitude to Dr. Giuseppe Amato and Dr. Claudio Gennaro, that offered me this opportunity, supported me, and advised me with patience during the research activities and the writing of this thesis. Without them, I would have not started nor ended this journey.

A special thank goes also to Dr. Fabrizio Falchi, with whom I collaborated closely in most of the research activities of my Ph.D. I am sincerely grateful for the invaluable help provided and the delightful discussions upon science, music, food, books, and much more is worth living for.

I gratefully acknowledge Prof. Francesco Marcelloni for supervising my Ph.D., and Prof. Michel Crucianu and Prof. Nicu Sebe that took the time to evaluate my work and provide valuable feedback.

I would like to thank the Networked Multimedia Information Systems lab for giving me the opportunity to join its team, especially Dr. Fausto Rabitti for the funding opportunities for trips and hardware.
I am sincerely grateful to all my lab mates for the three years spent together --- I really enjoyed my time at the CNR because of you. Special mentions go to Lucia Vadicamo, that helped me out countless times, Alejandro Moreo Fernàndez, for sharing ideas, music, and deep-learning jokes, Claudio Vairo and Franca Debole, for their collaboration and useful insights, Paolo Bolettieri, for the laughs about programming language master races, and Alessandro Nardi, for shooting some hoops together in the free time.

My deepest appreciation goes to all \emph{Gli Altri} with special mentions to Andrea, Giacomo, Giuditta, and Lorenzo, for the endless friendship and wonderful music we shared for more than 10 years, but also to Riccardo, Gianna, Sara, Lelle, Adele, and Lallu for providing an unlimited amount of good times and joy;
you are one of the pillars of my life, and I would not have gone so far without your support.

I am deeply grateful to my family: my parents Marta and Alessandro, that respectively taught me perseverance and criticism in their unique way, my sister Rita for constantly reminding me the important values in life without saying a word, and my uncle Paolo for introducing me to science in my youth;
you really helped to shape my existence in the best possible way.

Last but not least, incommensurable gratefulness is due to my girlfriend Regina, who stood next to me during the entire journey,
sharing ups and downs with me, enjoying with me every accepted publication, supporting me in deadline-dense stressful periods, helping me keeping my hair from falling while writing this thesis, loving me anyhow and anytime.
I owe you so much for all the invaluable time spent together that I will always treasure.