\chapter*{List of publications}

% for citing your paper, take directly APA format from Google scholar
% \item Surname, N., Surname, N. and Surname, N. (Year,Month). Title of the paper or journal. \emph{Place of publication}. (Vol. 123, pp. 456). EditorName.

\section*{International Journals}
\begin{enumerate}
    \item Amato, G., Carrara, F., Falchi, F., Gennaro, C., Meghini, C., and Vairo, C. (2017, April). Deep learning for decentralized parking lot occupancy detection. \emph{Expert Systems with Applications}. (Vol. 72, pp. 327-334). Pergamon.
    \item Carrara, F., Esuli, A., Fagni, T., Falchi, F., and Fernández, A. M. (2018, June). Picture it in your mind: Generating high level visual representations from textual descriptions. \emph{Information Retrieval Journal}. (Vol. 21, pp. 208-229). Springer.
    % manca mese, volume e pagine
    \item Carrara, F., Falchi, F., Caldelli, R., Amato, G., and Becarelli, R. (2018). Adversarial image detection in deep neural networks. \emph{Multimedia Tools and Applications}. (pp. 1-21). Springer.
\end{enumerate}

\section*{International Conferences/Workshops with Peer Review}
\begin{enumerate}
    \item Amato, G., Carrara, F., Falchi, F., Gennaro, C., and Vairo, C. (2016, June). Car parking occupancy detection using smart camera networks and deep learning. In \emph{2016 IEEE Symposium on Computers and Communication (ISCC)}. (pp. 1212-1217). IEEE.
    \item Carrara, F., Falchi, F., Caldelli, R., Amato, G., Fumarola, R., and Becarelli, R. (2017, June). Detecting adversarial example attacks to deep neural networks. In \emph{Proceedings of the 15th International Workshop on Content-Based Multimedia Indexing (CBMI)}. (p. 38). ACM.
    \item Amato, G., Carrara, F., Falchi, F., and Gennaro, C. (2017, June). Efficient Indexing of Regional Maximum Activations of Convolutions using Full-Text Search Engines. In \emph{Proceedings of the 2017 ACM on International Conference on Multimedia Retrieval (ICMR)}. (pp. 420-423). ACM.
    \item Vadicamo, L., Carrara, F., Cimino, A., Cresci, S., Dell’Orletta, F., Falchi, F., and Tesconi, M. (2017, October). Cross-Media Learning for Image Sentiment Analysis in the Wild. In \emph{2017 IEEE International Conference on Computer Vision (ICCV) Workshops}. (pp. 308-317).
    \item Amato, G., Bolettieri, P., Carrara, F., Falchi, F., Gennaro, C. (2018, June). Large-Scale Image Retrieval with Elasticsearch. In \emph{Proceedings of the 41st International ACM Conference on Research and Development in Information Retrieval (SIGIR)}. (pp.  925-928). ACM.
    \item Messina, N., Amato, G., Carrara, F., Falchi, F., and Gennaro, C. (2018, September). Learning Relationship-aware Visual Features. To appear in \emph{2018 IEEE European Conference on Computer Vision (ECCV) Workshops}.
    \item Carrara, F., Becarelli, R., Caldelli, R., Falchi, F. and Amato, G. (2018, September). Adversarial examples detection in features distance spaces. To appear in \emph{2018 IEEE European Conference on Computer Vision (ECCV) Workshops}.
\end{enumerate}

%\section*{Others}
%\begin{enumerate}
%    % for citing your paper, take directly APA format from Google scholar
%    \item Surname, N., Surname, N. and Surname, N. (Year,Month). Title of the paper or journal. \emph{Place of publication}. (Vol. 123, pp. 456). EditorName.
%\end{enumerate}
